---
title: 'Lecture 0: Basics of R programming'
author: "Xiao Guo"
date: "2023/3/11"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style type="text/css">

body, td {
   font-size: 14px;
}
code.r{
  font-size: 20px;
}
pre {
  font-size: 20px
}
</style>


# Overview

- ### Introduction to the R programming language
- ### Data Structures available in R: vectors, factors, matrices, lists, data frames
- ### Overview of conditional statements and functions
- ### Linear Regression Model

# 0.1 Introduction

### R is a programming language created and developed in 1991 by two statisticians at the University of Auckland, in New Zealand. It officially became free and open-source only in 1995. For its origins, it provides statistical and graphical techniques, linear and non-linear models, techniques for time series, and many other functionalities. Even if Python is the most common in the Data Science field, R is still widely used for specialized purposes, like in financial companies, research, and healthcare.

# 0.2 Requirements 

### If you want to start programming in R, you need to install the last versions of R and R studio. You are surely asking yourself why you need to install both. If you prefer, you can install only R and you will have a basic tool to write the code. In addition, R studio provides an intuitive and efficient graphical interface to write code in R. It allows to divide the interface into subwindows to visualize separately the code, the output of the variables, the plots, the environment, and many other features.

# 0.3 Assignments 

### When we program in R, the entities we work with are called objects. They can be numbers, strings, vectors, matrices, arrays, functions. So, any generic data structure is an object. The assignment operator is `<-`, which combines the characters `<` and `-`. We can visualize the output of the object by calling it:

```{r}
x <- 23
x
x <- 1/1+1*1
y <- x^4
z <- sqrt(y)
x
y
```

### You can use `sqrt, abs, sin, cos, tan, exp, and log` for basic mathematical operations. 

# 0.4 Vectors in R Programming

### In R, the vectors constitute the simplest data structure. The elements within the vector are all of the same types. To create a vector, we only need the function `c()`:

```{r}
v1 <- c(2,4,6,8)
v1
v2 <- 1:10
```

### The function `seq()` can also be applied to create more complex sequences. For example, we can add the argument by the step size and the length of the sequence:

```{r}
v2 <- seq(0,1,by=0.1)
v2
v3 <- seq(0,2,len=11)
v3
```

### To repeat the same number more times into a vector, the function rep() can be used:

```{r}
v4 <- rep(2,3)
v4
v5 <-c(1,rep(2,3),3)
v5
```

### There are not only numerical vectors. There are also logical vectors and character vectors:

```{r}
x <- 1:10
y <- 1:5 
l <- (x==y) 
l
cc <- c('a','b','c')
cc
```

# 0.5 Factors in R Programming

### Factors are specialized vectors used to group elements into categories. There are two types of factors: ordered and unordered. For example, we have the countries of five friends. We can create a factor using the function `factor()`.

```{r}
states <- c('italy','france','germany','germany','germany')
statesf<-factor(states)
statesf
```

### To check the levels of the factor, the function `levels()` can be applied.

```{r}
levels(statesf)
```

```{r}
a <- factor (c(1, 2, 3, 3))
a
levels(a)
```

# 0.6 Matrices in R Programming

### As you probably know, the matrix is a 2-dimensional array of numbers. It can be built using the function `matrix()`.

```{r}
m1 <- matrix(1:6,nrow=3)
m1
m1[1,]
m1[1:3,]
apply(m1, 2, sum)
m2 <- matrix(1:6,ncol=3)
m2
```

### It can also be interesting combine different vectors into a matrix row-wise or column-wise. This is possible with `rbind()` and `cbind()`:

```{r}
countries <- c('italy','france','germany')
age <- 25:27
rbind(countries,age)
countries <- c('italy','france','germany')
age <- 25:27
cbind(countries,age)
```

# 0.7 Arrays in R Programming

### Arrays are objects that can have one, two, or more dimensions. When the array is one-dimensional, it coincides with the vector. In the case it’s 2D, it’s like to use the matrix function. In other words, arrays are useful to build a data structure with more than 2 dimensions.

```{r}
a <- array(1:16,dim=c(6,3,2))
a
```

## lists

### The list is a ordered collection of objects. For example, it can a collection of vectors, matrices. Differently from vectors, the lists can contain values of different type. They can be build using the function `list()`:

```{r}
x <- 1:3
y <- c('a','b','c')
l <- list(x,y)
l
```

# 0.7 Data frames in R Programming

### A data frame is very similar to a matrix. It’s composed of rows and columns, where the columns are considered vectors. The most relevant difference is that it’s easier to filter and select elements. We can build manually the dataframe using the function `data.frame()`:

```{r}
countries <- c('italy','france','germany')
age <- 25:27
df <- data.frame(countries,age)
df
```

### R provides pre-loaded data using the function `data()`:

```{r}
data(mtcars)
head(mtcars)
```

### The function `head()` allows visualizing the first 6 rows of the mtcars dataset, which provides the data regarding fuel consumption and ten characteristics of 32 automobiles. 

### To check all the information about the dataset, you write this line of code:

```{r}
help(mtcars)
```

### To have an overview of the dataset’s structure, the function `str()` can allow having additional insights into the data:

```{r}
str(mtcars)
```

- ### The function `dim()` to look at the dimensions of the data frame
- ### The function `names()` to see the names of the variables

```{r}
dim(mtcars)
names(mtcars)
```

### The summary statistics of the variables can be obtained through the function `summary()`:

```{r}
summary(mtcars)
```

### We can access specific columns using the expression `namedataset$namevariable`. If we want to avoid specifying every time the name of the dataset, we need the function `attach()`:

```{r}
mtcars$mpg
attach(mtcars)
mpg
```

### We can also try to select the first row in the data frame using this syntax:

```{r}
mtcars[1,]
```

### Note that the index starts from 1, not from 0! If we want to extract the first columns, it can be done in this way:

```{r}
mtcars[,1]
```

### We can also try to filter the rows using a logical expression:

```{r}
mtcars[mpg>20,]
```

### We can also specify the column while we filter:

```{r}
mtcars[mpg>20,'mpg']
```

# 0.8 **for** and **while** in R Programming

### The **for** loop is used to iterate elements over the sequence like in Pandas. The difference is the addition of the parenthesis and curly brackets. It has slightly different syntax:

> ### for (var in seq) statement

```{r}
for (i in 1:4) 
{print(i)}
```

**while** executes a statement or more statements as long as the condition is true

> ### while (cond) statement

```{r}
i<-1
while (i<6)
{print(i)
  i<-i+1}
```

# 0.9 **if** statement in R Programming

### The syntax of the **if** statement is similar to the one in Python. As before, the difference is the addition of the parenthesis and curly brackets.

> ### if (cond1) {statement1} else {statement2}

> ### if (cond1) {statement1} else if {statement2} else {statement3}


```{r}
for (i in 1:4)
{if (i%%2==0) print('even') else print('odd')
}
```

### If we want to compare two numbers and see which number is greater of the other, we can do it in this way:

```{r}
a <- 10
b <- 2
if (b > a){
  print('b is greater than a')
}else if (a == b){
  print('a and b are equal')
}else {
  print('a is greater than b')
  }
```

### There is also a vectorized version of the if statement, the function `ifelse(condition,a,b)`. It’s the equivalent of writing:

> ### if condition {a} else {b}

```{r}
x<-3
ifelse(x>=0,'positive','negative')
```

# 0.10 Function in R Programming

### The function is a block of code used to perform an action. It runs only when the function is called. It usually needs parameters, that need to be passed, and returns an output as result. It’s defined with this syntax in R:

> ### namefunction <- function(par_1,par_2,…){
### A <- f(par_1,par_2)
### return(A) or list(outputname = A)
}

```{r}
average <- function(x)
{ val = 0
  for (i in x){val=val+i}
  av = val/length(x)
  av
    }
average(1:3)
```

# 0.11 Probability distributions in R Programming

### A characteristic of R is that it provides functions to calculate the density, distribution function, quantile function and random generation for different probability distributions. For example, let’s consider the normal distribution:

- ### `dnorm(x)` calculates the value of the density in x
- ### `pnorm(x)` calculates the value of the cumulative distribution function in x
- ### `qnorm(p)` calculates the quantile of level p
- ### `rnorm(n)` generates a sample from a standard normal distribution of n dimension













