---
title: 'Lecture 4: Data Manipulation'
author: "Xiao Guo"
date: "2023/3/14"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style type="text/css">

body, td {
   font-size: 14px;
}
code.r{
  font-size: 20px;
}
pre {
  font-size: 20px
}
</style>

# 4.1. Data Manipulation

## Data Exploration

- ###  R for Data Science (O'Reilly 2017) by Hadley Wickham

- ### Free Online: http://r4ds.had.co.nz/

- ### Major coverage:
  -  #### Data manipulation (R:dplyr)
  -  #### Data visualization (R:ggplot2)

- ### Data Exploration = Data manipulation + Data visualization.


## Data Wrangling

- ###  Data wrangling (a term used in data science) is the process of transforming/mapping data from raw format into ready-to-analyze format.

- ### Besides ggplot2() for data visualization, Hadley Wickham has created a series of R packages for data wrangling, including

  - #### tidyr for tidy data: observations in rows, variables in columns
  - #### tibble for better ways to create, print and subset data frames
  - #### dplyr for data manipulation -> today

##  Feature Engineering

- ### A term often used in machine learning
> #### Andrew Ng (Stanford): "Coming up with features is difficult, time-consuming, requires expert knowledge. Applied machine learning is basically feature engineering."

- ### In statistics: variable creation and transformation

- ### Dictionary learning with overcomplete features ...

- ### Nowadays, deep learning algorithms aim at automatic feature learning instead of manual feature engineering ...

# 4.2 R:dplyr Package

## R::dplyr verbs

- ### `filter()` to select observations
- ### `arrange()` to order observations
- ### `mutate()` to add new variables
- ### `group_by()` to group variables for summarise
- ### R::base:`merge()` to combine two data.frames (or R::dplyr xxx_joins)

## Filter

```{r}
library(dplyr)
(tmp = filter(iris, Species == 'versicolor' & Sepal.Length > 6.6))
```

- ### Rowwise selection of samples/observations
- ### Similar to base:which or subsetting

## Arrange

```{r}
arrange(tmp, Sepal.Length, Sepal.Width, desc(Petal.Length))
```

- ### Similar to base:sort and order functions

## Mutate 

```{r}
tmp = mutate(iris, Sepal.Size = Sepal.Length*Sepal.Width, 
             Petal.Size = Petal.Length*Petal.Width) 
head(tmp)
```

- ### New variable/feature creation
- ### Base commands: tmp$Sepal.Size = ...

```{r}
library(ggplot2)
ggplot(tmp, aes(x=Sepal.Size, y=Petal.Size, colour=Species)) + 
  geom_point(size=2)
```

## Summarise
```{r}
summarise(group_by(tmp, Species), mean(Sepal.Size), mean(Petal.Size))
```

## Merge
```{r}
(tmp1 = data.frame(Species=levels(iris$Species), x1 = c("A","B","C"), x2 = round(runif(3),3)))
head(merge(iris, tmp1, by = "Species"))
```

# 4.3 Pipes %>%

- ### The pipe %>% requires R package dplyr or magrittr

- ### Powerful trick for coding a sequence of operations

- ### Output of old operation as the first argument of new operation

- ### Especially useful in combined with ggplot2

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(mean_mpg = mean(mpg))
```







